Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-02-16 draft.tex

@@

Despite the reasonable fit of the evolving population model, the fit to tardy SNe diverges from the data at high stellar mass, corresponding to the oldest average stellar age. It is difficult to reconcile this turnover with the simple DTD models used thus far in this work, as it would require a steepening or complete turn-off of the DTD at late ($\tau\gtrsim 5$~Gyr) times. Here we describe two possible ways in which the model can be reconciled with the data in the highest-mass galaxies.

One possibility is that sub-luminous SNe are more numerous in such predominantly old stellar populations, such that the fraction passing our light curve cuts is lower than the rest of the sample. Such a phenomenon could be caused by SN Ia sub-classes such as SN 1991bg-like SNe, which are known to explode exclusively in old environments \citep{Perets2010} at large delay times $>6$~Gyr \citep{Panther2019}.

>>>>

The SN Ia rates generated by the posterior predictive checks are shown in Fig. \ref{fig:rate_fitted_split_x1_joint}. The best fitting values are $t_1 = 0.65 \pm 0.1$~Gyr and $t_2 = 0.85 \pm 0.15$~Gyr, corresponding to a relatively sharp transition between the two populations.

@@

Despite the reasonable fit of the evolving population model, the fit to tardy SNe diverges from the data at high stellar mass, corresponding to the oldest average stellar age. It is difficult to reconcile this turnover with the simple DTD models used thus far in this work, as it would require a steepening or complete turn-off of the DTD at late ($\tau\gtrsim 5$~Gyr) times.

One possibility is that sub-luminous SNe are more numerous in such predominantly old stellar populations, such that the fraction passing our light curve cuts is lower than the rest of the sample. Such a phenomenon could be caused by SN Ia sub-classes such as SN 1991bg-like SNe, which are known to explode exclusively in old environments \citep{Perets2010} at delay times as long as $>6$~Gyr \citep{Panther2019}. If the SN 1991bg-like objects are drawn from the same parent population of WDs as the "normal" SNe Ia, they could explain the lack thereof in the sample.
